
  % How generation interval inference is done in practice

  % There are essentially two tiers of methods:

  % 1. Population-level / naive pooling (most common). All observed generation intervals (or serial
  % intervals) are treated as iid draws from a parametric distribution, typically Gamma. Parameters are
  %  estimated by MLE. This is the "population-level method" in Park et al. 2020 (Section 7.4):
  % estimate the mean and shape of a Gamma from the aggregated data, optionally correcting for
  % right-truncation by exponential reweighting (their Eq. 7.7). This approach is used by the majority
  % of empirical studies — for example, the widely-cited COVID-19 generation interval estimates from
  % Ganyani et al. and others.

  % 2. Individual-level Poisson process models (more sophisticated). Each infector j's transmission is
  % modeled as a non-homogeneous Poisson process with rate proportional to the GI density. This is Park
  %  et al.'s Eq. 7.8–7.11. The likelihood for infector j with n_j offspring at generation intervals
  % s_{i,j} is:

  % L_j ∝ Λ^{n_j} exp(−Λ F₀(t_trunc − t_j)) × ∏ᵢ f₀(s_{i,j})

  % This accounts for right-truncation and the number of offspring. Hart et al. (2022, eLife) extend
  % this further using data-augmentation MCMC for household data, jointly estimating infection times
  % and transmission trees.

  % The critical point: in both approaches, generation intervals from the same infector are treated as
  % conditionally independent. Park et al. acknowledge this explicitly after their Eq. 7.11:

  % "In theory, the forward generation intervals arising from the same infector may be correlated
  % because of non-independence in the contact process [35]; although we do not account for this
  % potential correlation in our likelihood..."

  % They note this is a known gap but find empirically that the iid approximation works adequately for
  % their purposes (estimating R₀ from r).

  % No published method I found accounts for within-infector correlation of the kind your model
  % introduces. The correlation Park et al. reference (from the contact process) is a network effect.
  % Your model introduces a more fundamental source: the shared latent component l_i from the
  % infectiousness profile itself.

  %   Sources:
  % - https://royalsocietypublishing.org/doi/10.1098/rsif.2019.0719
  % - https://elifesciences.org/articles/70767
  % - https://royalsocietypublishing.org/doi/10.1098/rspb.2015.2026
  % - https://royalsocietypublishing.org/doi/10.1098/rsif.2020.0756
  % - https://journals.plos.org/ploscompbiol/article?id=10.1371%2Fjournal.pcbi.1014239
  % - https://pmc.ncbi.nlm.nih.gov/articles/PMC10990235/






















  \clearpage
\section{Discussion}

\begin{itemize}
  \item Brief summary of what we did 
  \item High level: the share of within- vs. between-individual variation in infection timing has a substantial impact on both uncontrolled and controlled epidemics. 
  \item To illustrate this, we introduce a novel modeling framework, the gamma time-shifted model. 
  \item We combine this with a general contact process model that extends to the individual level the basic renewal equation epidemic model. 
  \item We expect this will have implications for gathering contact tracing data and for building projection and forecasting models, highlighting a key point of underexplored uncertainty and setting forth a way of dealing with it parsimoniously, even if just in sensitivity analyses. 
\end{itemize}

\begin{itemize}
  \item Why have we gone so long without really caring about $\psi$? One: it doesn't impact mean-field dynamics, by design. We can get good assessments of the overall behavior of an epidemic without it. Also, we've dealt with it implicitly a lot, we just haven't isolated it as an independent axis of variation
  \item For example: individual-based models make assumptions about $\psi$ all the time, as  do the papers we referenced in the second-to-last paragraph of the intro. 
  \item We can also adjust $\psi$ using methods like the linear chain trick, stringing together sets of dummy exposed and infectious categories; but this also changes the generation interval distribution. Figuring out how to vary the individual ifnectiousness profile without affecting the population-level generation interval distribution is the main contribution of this work. 
\end{itemize}

\begin{itemize}
  \item For the time shift story: the shift in $E[\varsigma]$ is minor; the inflation of $\text{Var}[\varsigma]$ is the real story. 
  \item For the inference of generation intervals and r: this suggets we need to develop good statistical tools and we need to be careful as we collect our data. 
  \item for the overdispersion bit: this a fundamentally new mechanism of superspreading that we should be aware of. A key point: ``super-shedders' and ``super-contacters'' are theoretically identifiable as individuals, but when superspreading arises due to poorly timed infectiousness, that's not necessarily something we can anticipate. It suggests there might be a limit to which superspreading can be targeted, and opens up the need for us to bettter understand the relative contributions of these mechanisms to overdispersion. 
  \end{itemize

\begin{itemize}
  \item  For the sensitivity of TE: note that shifts in symptom timing are commonplace (happend with SARS-CoV-2), and small changes in testing frequency are easy operational decisions. So, understanding the shape of the individual infectiousness profile is important for anticipating how these small shifts might impact epidemic behavior. Also, something we didn't explore here but could be fleshed out more is that we might even use these kinds of shifts, or maybe testing, to assess the shape of the individual infectiousness profile -- they might give us information beyond just simple contact tracing data. 
  \item More details on the counter-intuitive impact of TE/GE contraction/overdispersion for detect-and-isolate, and Reff/overdispersion for gathering size restrictions. We often think of interventions in terms of how they reduce $R_{eff}$, but that's a bit dangerous, because the full effect can work in multiple directions (reduction in $R_{eff}$ but not a commensurate drop in epidemic growth rate, for example). The framework here gives a robust way of dealing with this beyond just naive reductions in R. 
\end{itemize}

\begin{itemize}
  \item Identifiability is a challenge, but seeems theoretically possible under reasonably good circumstances. 
  \item Available data suggests that there may be a spectrum of $\psi$-values across pathogens, though there are huge limitations to the data and the inference methods we've used here, which we discuss in a moment. 
  \item That said, there are reasons to believe there could be differences in $\psi$ across pathogens; some we know generate long-term shedding while others might be short and fast, and our findings roughly align with this taxonomy. 
  \item There are local-level impacts that could influence the observed $\psi$, like local susceptible depletion: for measles, maybe we have a short infectiousness window because everybody gets infected immediately. In theory, a person could go on infecting, but they've already burned through  all their contacts as soon as they reach the infectious stage. That's one thing that could happen and is hard to disentangle with existing data. (mabye epidemiological challenge studies, like when they stuck hamsters in the air vents of people with tuberculosis?) 
\end{itemize}

\begin{itemize} 
  \item Limitations abound: the gamma time shift model is just one choice 
  \item the contact process model is hard to ground empirically 
  \item Lots more could be done with test-based control assessments 
  \item inference is wayyyyy over-simplified; we should talk in detail about all the various confounders that could imapct this, many of which are in the comments at the end of code/psi_empirical_litgi.R. 
\end{itemize }

\begin{itemize}
  \item But, we still think this will have an impact. It opens up new avenues of research and highlights a fundamental epidemiological quantitiy worth studying in its own right. 
  \item We need better data, and better ways of understanding how to use existing data to get at $\psi$. 
  \item We expect this will help generate better projections and forecasts -- or at least projections and forecasts that are more honest about unertainty, and clearar about a potential mechanism driving some of that uncertainty. Thus, we hope it will impact public health practice overall. 
\end{itemize}



























  ---
  1. Benchmark pathogen parameters
  - Table of $R_0$, $\bar{g}$, $\text{Var}[g]$, $\alpha$, $\beta$ for influenza, SARS-CoV-2 omicron,
  and measles. (Referenced as "Table XX" in results.tex line 62 and model.tex line 93; already
  appears in supplementary.tex but needs a main-text version or cross-reference.)
  - Literature sources for each pathogen's parameters.

  2. Simulation details
  - Already summarized in model.tex (simulation approach subsection), but results.tex refers to
  "Methods" for additional details in several places. The current model.tex section points to
  "Supplementary Methods" for full details. Methods should at minimum cross-reference or briefly
  restate the key simulation parameters: $N = 10{,}000$, $n_\text{sim} = 5{,}000$, epidemic
  establishment threshold of 500 infections.

  3. Growth rate estimation procedure
  - Results.tex line 80: "we fit a Poisson generalized linear model to the daily infection counts."
  Methods should specify: infinite population, 7 days of data after 100 cumulative cases, daily
  binning, Poisson GLM with log link and day as covariate.

  4. Generation interval estimation procedure
  - Results.tex lines 82–93: Simulated offspring from index cases, applied Gamma observation noise
  ($a_\text{obs}$, $b_\text{obs}$), extracted serial intervals, estimated posterior for $(\alpha,
  \beta)$ treating observations as independent. Methods should specify: number of index cases (100 in
   results.tex line 93, but 5,000 in uncontrolled.tex line 80), the posterior estimation approach
  (what prior, what likelihood), the coverage calculation (95% credible interval containing the true
  value), and the observation delay parameters ($a_\text{obs} = 4$, $b_\text{obs} = 1$ from
  results.tex line 83).

  5. ICC and effective sample size (Kish's design effect)
  - Results.tex lines 82–89: The ICC formula and $n_\text{eff} = n / (1 + (R_0 - 1) \cdot
  \text{ICC})$. Methods (or supplementary methods) should derive this. The ICC formula itself is
  stated in results.tex but the derivation is promised as "Supplementary Materials."

  6. Overdispersion estimation from simulations
  - Results.tex line 121: "We estimated OD by moment-matching the mean and variance of the offspring
  counts to a Negative Binomial distribution." Methods should state this explicitly.
  - Also specify: 10,000 index cases per scenario, range of $c_\text{amp}$, $\sigma$, $\psi$ values
  swept.

  7. Testing effectiveness (TE) numerical integration
  - Results.tex line 141: "TE can then be calculated via numerical integration (Methods)." This is
  explicitly promised.
  - Controlled.tex lines 19–21 give the integral: $\text{TE} = \eta \rho \int_{-\infty}^\infty (1 -
  F_\epsilon(t)) f_\text{det}(t) , dt$. Methods should specify:
    - The two detection mechanisms (symptom-based and test-based)
    - Symptom-based: $D \sim \text{Normal}(\delta_\text{symp}, \sigma_\text{symp}^2)$ relative to
  peak infectiousness, immediate isolation
    - Test-based: tests at intervals $\Delta_\text{test}$, detectability window $[\delta_\text{pre},
  \delta_\text{post}]$ around peak, $\text{Exp}(\lambda_\text{act})$ turnaround delay
    - How $f_\text{det}$ is constructed for each mechanism
    - Default parameter values used in figures

  8. Detect-and-isolate simulation details
  - Results.tex line 167: Simulation parameters for D&I epidemic simulations ($\rho = 0.5$?, $\eta =
  1$?, $\delta_\text{symp}$, $\sigma_\text{symp}$, $\Delta_\text{test}$, etc.). The controlled.tex
  draft (line 42) specifies $\rho = 0.5$, $\eta = 1$, deterministic isolation 1 day before peak.
  Methods should state these and the comparison setup (naive $R_\text{eff}$-matched simulations).

  9. Generation interval truncation formula
  - Results.tex lines 151–159: $g^*(\tau)$ formula. The derivation is promised as "Supplementary
  Methods."

  10. $\nu_i^*$ under D&I and overdispersion
  - Results.tex lines 161–165: $\nu_i^* = \nu_i F_\epsilon(m_\epsilon + \iota_i)$. Methods or
  supplementary should derive this and the associated dispersion index expressions.

  11. Gathering size restriction details
  - Results.tex lines 174–188: The truncated stochastic contact model, $R_\text{eff} = \mu_T R_0$, OD
   formulas. Methods should specify the truncation mechanism, the values of $\sigma$ and
  $c_\text{max}$ used, and the three-way comparison setup (truncated vs. $R_\text{eff}$-adjusted
  stochastic vs. $R_\text{eff}$-adjusted uniform).

  12. Psi identifiability: simulation-based power calculation
  - Results.tex lines 196–198: 10/25/50/100 index cases, $\psi \in {0, 0.5, 1}$, three pathogens,
  Gamma observation delays (fast/medium/slow), flat prior on $\psi$, 200 replicates, power defined as
   >80% of simulations with 95% posterior mass within 0.2 of true $\psi$. Methods should detail:
    - The posterior computation (likelihood, prior, grid)
    - The power criterion
    - The observation delay parameters (mean 1, 2, 4 days)

  13. Empirical $\psi$ estimation
  - Results.tex lines 209–211 (your new section): Methods should describe:
    - Data extraction from OutbreakTrees database
    - Supplementary COVID-19 data sources (Ganyani, Hart)
    - Cluster formation (grouping by index case, restricting to $\geq 2$ secondaries)
    - Likelihood: integration over shared $l_i$, product of SI densities within cluster
    - Prior: flat on $\psi \in [0, 1]$, grid of 101 points
    - Fixed GI parameters from literature (and their sources, per pathogen)
    - Incubation period parameters used
    - Summary statistics: posterior mode, mean, 95% CI

  14. Items from the older draft that the results reference as "Supplementary Methods"
  These are derivations already partly present in supplementary.tex but should be cross-referenced
  from Methods:
  - Deriving population-level quantities from individual-level components (supplementary.tex line 4)
  - $\text{Var}[W]$ derivation and the Gamma approximation for $W|\text{surv}$ (already in
  supplementary.tex)
  - Overdispersion in daily infection counts (already in supplementary.tex)
  - OD formula for periodic contacts (already in supplementary.tex)
  - OD formula for stochastic contacts (already in supplementary.tex)
  - TE integration details
  - $g^*(\tau)$ derivation
  - Gathering size restriction OD formulas

  15. Software and reproducibility
  - R version, key packages (odin for simulations, tidyverse, etc.)
  - Code availability statement






















    ---
  The core parallel to Lloyd-Smith and where it breaks down

  Lloyd-Smith 2005 works because it has a ruthlessly simple structure: one new parameter (k), one
  empirical finding (k is small for many diseases), and one punch line (individual-specific control
  beats population-wide control). The paper is 4 pages. Every figure advances the same argument. The
  reader walks away knowing exactly what k is, why it matters, and what to do about it.

  Your paper has the same potential with psi, but the current structure doesn't deliver it with the
  same force. Here are the specific issues:

  ---
  1. The paper has too many results and no clear headline

  You have seven results subsections, each presenting a different consequence of psi. This breadth is
   intellectually impressive but strategically diluting. The reader is left with "psi affects a lot
  of things" rather than one crystallized takeaway. Lloyd-Smith's Figure 1 puts the empirical
  evidence front and center on page 1; by the end of page 2 you know superspreading is the norm. Your
   most novel finding --- that psi generates a new mechanism of superspreading through the g_i-c_i
  interaction (Section 3) --- is buried as the fourth results subsection, after time shifts and
  inference, which are less paradigm-shifting.

  Consider reorganizing the results around a hierarchy:
  - Lead with the new superspreading mechanism. This is the finding that will make modelers sit up.
  It's genuinely new, it connects directly to Lloyd-Smith's framework (psi generates overdispersion
  in offspring counts through a mechanism invisible to the existing k framework), and it has a clean
  analytic expression. If you want psi to play the role that k plays, you need to show early that psi
   creates overdispersion --- not just that it inflates variance in other quantities.
  - Then show the empirical estimates. Lloyd-Smith puts data in Figure 1. Your empirical psi
  estimates across 8 pathogens (currently the last results subsection) should come much earlier,
  ideally right after establishing why psi matters. This gives the reader concrete numbers to anchor
  to: "psi varies widely, and for some diseases it's near zero --- which means these effects are
  real, not hypothetical."
  - Then cover interventions (D&I and gathering restrictions), which are the practical implications.
  - Relegate time shifts and inference to supporting roles --- either later in the results or
  partially in supplementary. These are interesting but don't carry the same weight as "there's a new
   source of superspreading your models are missing."

  2. The first two results subsections read as model development, not findings

  Section 1 ("A time-shifted Gamma model flexibly captures...") and Section 2 ("Expressing the
  individual reproduction number...") define the model and notation. These aren't results --- they're
   framework. In Lloyd-Smith, the model (negative binomial offspring distribution) is introduced in a
   single paragraph of the main text. Your model is more complex, but the exposition could be
  tightened. Consider moving the model definition to a brief "Model" subsection before Results, or
  folding it into Methods. The current structure means the reader doesn't encounter an actual finding
   until the third subsection (page ~4-5 of results), which is too late.

  3. The model.tex section creates redundancy

  You have a "Modeling framework" section (model.tex) that overlaps substantially with Results
  subsections 1-2 --- both define the Gamma time-shift model, the contact process, and the simulation
   approach. This will confuse readers. Pick one location for the model definition and commit to it.

  4. The D&I section tries to do too much

  Section 5 covers three distinct phenomena --- TE sensitivity, generation interval truncation, and
  overdispersion from isolation --- each of which is complex. Consider whether you need all three at
  equal depth. The TE sensitivity is the most intuitive and actionable; the truncation and
  overdispersion results could be condensed or split into a separate subsection.

  5. The paper needs a sharper "so what" for non-modelers

  Lloyd-Smith ends with concrete, actionable proposals: report p_0 (proportion non-transmitting),
  define SSEs using the Z^(n) threshold, target the most infectious 20%. These are things
  epidemiologists can immediately do. Your discussion mentions "incorporating psi even as a
  sensitivity parameter," but this isn't crystallized into a clear call to action.

  Some concrete recommendations you could make:
  - Contact tracing studies should record index case identity for each serial interval (you mention
  this, but it could be elevated)
  - Report within-cluster variance of serial intervals alongside the mean serial interval
  - Sensitivity analyses in projection models should sweep psi in [0, 1]
  - When evaluating D&I interventions, assess sensitivity of TE to detection timing shifts

  6. The title is descriptive rather than provocative

  "The impact of punctuated infectiousness on infectious disease dynamics" tells the reader what the
  paper is about but not what it finds. Compare Lloyd-Smith: "Superspreading and the effect of
  individual variation on disease emergence." Consider something that conveys the finding, e.g.:
  - "Individual infectiousness timing as a hidden driver of epidemic variability and superspreading"
  - "A missing axis of variation in epidemic models: the shape of individual infectiousness"

  7. The abstract is missing

  This is critical for framing. For maximum impact, the abstract should lead with the gap (we model
  A(tau) but ignore a_i(tau)), state the key finding (narrow a_i(tau) generates overdispersion
  through a new mechanism, and makes interventions sensitive to timing), give the empirical anchor
  (psi varies widely across pathogens), and end with the implication.

  8. Figure strategy

  You need a "hero figure" --- the equivalent of Lloyd-Smith's Figure 1. Currently your figures are
  all simulation-output panels. Consider a composite figure that puts the conceptual schematic (your
  Figure 1, which is good), the empirical psi estimates (currently Figure 7), and one key consequence
   (e.g., the OD heatmap) side by side. This would let the reader see "here's the concept, here's the
   data, here's why it matters" in a single visual.

  Also: all your figure captions are "[Caption]". This matters more than you might think --- strong
  captions that tell a story independently of the main text are a hallmark of high-impact papers.

  9. For the Nature vs. JRSI decision

  If you target Nature:
  - You need to cut to ~3,000 words of main text, with everything else in Supplementary
  - Lead with the superspreading mechanism + empirical data
  - The interventions section becomes a brief third act
  - Time shifts and inference move to supplementary
  - You need the hero figure

  If you target JRSI:
  - You can keep the current breadth
  - But you still benefit from reorganizing to lead with the strongest result
  - The longer format lets you develop the D&I and gathering size results fully
  - The comprehensive supplementary derivations are a natural fit

  My honest assessment: the ideas are strong enough for Nature. The question is whether you can
  distill them to the required length without losing the story. The superspreading mechanism +
  empirical estimates + intervention sensitivity, compressed to 4 pages with a killer figure, could
  work. But the current structure, with its seven results subsections and dual model definitions,
  would need significant restructuring to fit that format. If the restructuring feels forced, JRSI is
   a strong venue where the full breadth would be appreciated and influential among exactly the
  modeler audience you're targeting. Champredon & Dushoff (2015) in Proc. R. Soc. B is another
  comparable precedent --- well-cited, influential in the modeling community.

  10. One more thing: the language around psi

  You consistently use "punctuated" and "smooth" to describe the extremes. This is clear but slightly
   technical. Lloyd-Smith's language --- "superspreaders," "the 20/80 rule" --- immediately connects
  to intuition. Consider whether there's a more vivid way to describe what low psi means in practice:
   "burst transmitters" vs. "sustained shedders," or "synchronized" vs. "diffuse" infectiousness. A
  memorable label helps a parameter become part of the field's vocabulary.



































  ❯ Cool. Do you think I should focus more on targeting a nature-style journal (maybe also
  science), or a proceedings b/royal society interface journal? What should be my manuscript
  submission strategy?

⏺ This depends on what you're optimizing for, and I want to be honest about the trade-offs rather
  than just cheerleading.

  The case for Nature/Science:

  The conceptual contribution is genuinely at the right level. You're introducing a new parameter
  that captures a previously unrecognized axis of variation, you're showing it generates a novel
  mechanism for a phenomenon (superspreading) that the field cares deeply about, and you have
  empirical estimates across multiple pathogens. That's structurally parallel to Lloyd-Smith 2005.
  The "new mechanism of superspreading" angle is the kind of finding that Nature editors respond to
  --- it's surprising, general, and changes how you think about a well-studied phenomenon.

  The case against Nature/Science:

  A few things work against you relative to Lloyd-Smith's position in 2005. First, timing:
  Lloyd-Smith published during the immediate aftermath of SARS-1, when superspreading was an urgent,
  unresolved public health puzzle. Your paper doesn't have that kind of topical urgency --- it's a
  fundamental contribution to epidemic theory, but there's no ongoing crisis creating demand for it.
  Nature and Science weight timeliness heavily.

  Second, the effects of psi, while real and theoretically important, are effects on stochastic
  dynamics --- they're invisible to deterministic models by construction. This is intellectually
  interesting but makes the "so what" harder to communicate to a broad audience. Lloyd-Smith's k had
  a visceral, immediate interpretation: 20% of cases cause 80% of transmission, and you can see it in
   the data. Your strongest analogue is the OD formula, which is elegant, but the practical
  consequences (higher extinction probability, sensitivity of TE to detection timing) require more
  work from the reader.

  Third, the empirical estimates are preliminary by your own admission. Lloyd-Smith had clean
  maximum-likelihood fits to outbreak data that decisively rejected the Poisson model. Your estimates
   are Bayesian posteriors with fixed GI parameters and acknowledged ascertainment biases, and the
  most striking results (measles and plague at psi near 0) are the ones most susceptible to those
  biases. A Nature referee will press hard on this.

  My honest assessment:

  I think the ideas are strong enough that a Nature/Science submission is defensible, but I'd put the
   probability of acceptance at maybe 10-15% --- which is roughly the acceptance rate for papers that
   make it past the editors' desk. The risk is a 6-12 month cycle of submission, desk rejection or
  review, revision, and possible rejection, after which you submit to Proc B / Interface anyway.

  What I'd actually recommend:

  Submit to Nature first, but with a short time horizon. Nature's initial editorial decision (desk
  accept/reject) typically comes within 1-2 weeks. If it passes the editors, great --- you're in
  play. If it's desk-rejected, you've lost almost no time. You could also try Science, but I'd do one
   or the other, not both sequentially --- the marginal value of the second attempt isn't worth the
  delay.

  If Nature desk-rejects, go to PNAS or Proc B next, not Interface. Here's why:

  - PNAS (via direct submission or contributed) has the prestige and readership to make psi part of
  the field's vocabulary, accepts longer papers than Nature, and publishes a lot of mathematical
  epidemiology. The format would let you keep the full argument without the compression pain of a
  Nature Letter. It's also faster than Interface in my experience.
  - Proc B is the natural home for this kind of work --- the Champredon & Dushoff paper is there, as
  are many of the papers you cite. It's well-read by exactly the modelling community you're
  targeting. The format fits your content well.
  - Interface is more methods-oriented. Your paper has substantial methodological content, but the
  story is really about epidemiological insight, not methodology. Proc B is a better fit.

  So my suggested sequence would be:

  1. Nature (1-2 weeks for desk decision; low cost if rejected)
  2. If desk-rejected: PNAS or Proc B, depending on whether you want to optimize for breadth of
  readership (PNAS) or depth within the modelling community (Proc B)

  One practical consideration: Nature requires ~3,000 words; Proc B allows ~5,000-6,000. You already
  have both drafts. If Nature desk-rejects, you can submit the Proc B version within days rather than
   rewriting from scratch. Having both ready is a genuine strategic advantage --- it means a desk
  rejection doesn't cost you more than two weeks.

  One thing I'd do regardless of venue: write a strong cover letter that explicitly makes the
  Lloyd-Smith parallel. "Just as Lloyd-Smith et al. (2005) showed that variation in the amount of
  individual transmission (parameterized by k) has profound consequences for epidemic dynamics, we
  show that variation in its timing (parameterized by psi) has consequences of comparable magnitude,
  through mechanisms that are invisible to existing frameworks." That framing helps editors see where
   the paper sits in the field's intellectual history.

✻ Worked for 40s

❯ Could you draft a nature-style cover letter and put it in the writeup/v4_nature directory?

⏺ Write(cover_letter.tex)
  ⎿  Wrote 55 lines to cover_letter.tex
     \documentclass[11pt]{letter}

     \usepackage[margin=1in]{geometry}
     \usepackage[T1]{fontenc}
     \usepackage{mathpazo}
     \usepackage{hyperref}
     \usepackage{xcolor}

     \signature{Stephen M.~Kissler, PhD\\
     Assistant Professor\\
     … +45 lines (ctrl+o to expand)

⏺ Bash(cd "/Users/stephen_ally/Dropbox
      (Personal)/DropboxHarvard/Projects/IndividualInfectiousnessProfiles/writeup/v4_nature" &&
      pdflatex -interaction=nonstopmode cove…)
  ⎿  Error: Exit code 1

⏺ Bash(cd "/Users/stephen_ally/Dropbox
      (Personal)/DropboxHarvard/Projects/IndividualInfectiousnessProfiles/writeup/v4_nature" &&
      pdflatex -interaction=nonstopmode cove…)
  ⎿  ! Undefined control sequence.
     l.36 ... interventions purely in terms of $R_\text
                                                       {eff}$ can be misleading, ...
     … +12 lines (ctrl+o to expand)

⏺ The \text command needs amsmath. Let me fix that.

⏺ Update(cover_letter.tex)
  ⎿  Added 1 line
       4  \usepackage[T1]{fontenc}
       5  \usepackage{mathpazo}
       6  \usepackage{hyperref}
       7 +\usepackage{amsmath}
       8  \usepackage{xcolor}
       9
      10  \signature{Stephen M.~Kissler, PhD\\

⏺ Bash(cd "/Users/stephen_ally/Dropbox
      (Personal)/DropboxHarvard/Projects/IndividualInfectiousnessProfiles/writeup/v4_nature" &&
      pdflatex -interaction=nonstopmode cove…)
  ⎿  Done: 2 pages

⏺ Read 1 file (ctrl+o to expand)

⏺ Looks clean. The letter runs about 1 page of substance, which is appropriate for Nature --- editors
   don't read long cover letters.

  A few things you may want to adjust before sending:

  - Reviewer suggestions: You'll want to confirm these are people you're comfortable having review
  the paper. Lloyd-Smith and Dushoff are the most natural choices given the intellectual lineage, but
   if either has a close connection to you or your co-authors, swap them out. Fraser is a strong
  choice for the D&I angle. You might also consider Adam Kucharski (LSHTM) for the
  superspreading/overdispersion angle, or Joel Miller for the branching process theory.
  - The date will auto-generate from compilation time, so it'll be correct whenever you submit.
  - Co-author approval: The letter states all authors have approved --- make sure that's true before
  submitting.
  - Competing interests / funding: Both are declared as none / to be completed. Fill these in before
  submission.

✻ Worked for 1m 50s

❯ Ok, now: with the nature submission, there are a few key weaknesses I'm concerned about in
  addition to the ones you mentioned above. First: we say that weekly variation is the most
  impactful for influenza/sars-cov-2/measles, but we don't show that there's meaningful weekly
  variation in actual contact patterns. Is that a critical piece of the story? Second: can we
  really call this a "new mechanism of superspreading?" Maybe it's because I"m so close to the
  work now, but I worry a reviewer will read our description that "when the infectiousness
  window coincides with a period of high contacts, we get superspreading" and say, effectively,
  duh, that's trivial. Those are the main ones: anything i should do about them? does it tank
  nature/science as a venue?

⏺ These are both legitimate concerns. Let me take them seriously.

  On weekly contact variation:

  This is a real gap, but I don't think it's fatal. The claim you're making isn't "weekly contact
  variation exists and causes superspreading via psi." The claim is "when contact rates vary on
  timescales comparable to the mean generation time, psi determines how much overdispersion results."
   The weekly example is illustrative, not the finding itself. The finding is the formula (Eq. 1) and
   the mechanism it captures.

  That said, a referee will reasonably ask: "Is there evidence that contacts vary on the relevant
  timescales?" And the answer is yes --- there's a substantial literature on this. Weekend vs.
  weekday contact differences are well-documented in contact survey data (POLYMOD, CoMix, BBC
  Pandemic study). These studies consistently show 20-40% fewer contacts on weekends than weekdays,
  which is roughly c_amp in the range 0.1-0.2. There's also variation at longer timescales (school
  terms, holidays) and shorter ones (daily routines). You don't need to generate new data on this;
  you need to cite it. A paragraph in the Discussion or Methods saying "empirical contact surveys
  document substantial day-of-week variation (refs), with weekend contact rates 20-40% lower than
  weekday rates, corresponding to c_amp ~ 0.1-0.2 at c_per = 7 days" would close this gap. You could
  even plug those empirical values into the OD formula to give concrete estimates of psi-attributable
   overdispersion, which would strengthen the paper.

  What I wouldn't do is add a new analysis of contact data. That's scope creep, and it's not where
  the contribution lies.

  On "is this really a new mechanism?":

  This is the more serious concern, and I want to be careful here because I think the answer is
  nuanced.

  A referee who says "duh, if your infectious window coincides with high contacts you transmit more"
  is correct at the level of the individual. That observation is not new. What's new is the
  population-level consequence and the specific conditions under which it matters. There are three
  things the paper shows that are genuinely non-obvious:

  First, this mechanism generates overdispersion even when there is zero heterogeneity in biological
  infectiousness and zero heterogeneity in average contact rates. In the standard Lloyd-Smith
  framework, you need variation in nu_i to get overdispersion. You're showing that identical
  individuals with identical average contacts can still produce overdispersed offspring counts,
  purely through timing. That's not "duh" --- it's a specific, falsifiable claim that contradicts the
   implicit assumption of most superspreading analyses.

  Second, the magnitude of the effect is governed by a precise relationship between psi, the
  generation time, and the contact period (the OD formula). The fact that this has a clean analytic
  form, and that it predicts when the effect is large vs. negligible, is not something you'd get from
   the intuition alone.

  Third, and this is the key point that I think you should emphasize more: this overdispersion cannot
   be targeted. Classical superspreading from super-shedders or super-contactors is in principle
  addressable by identifying and isolating high-risk individuals. The timing-contact overdispersion
  cannot be, because it doesn't arise from any persistent individual trait. This has a concrete
  policy implication: it sets a floor on the overdispersion that remains even after you've perfectly
  identified and controlled all super-shedders and super-contactors.

  The "duh" objection mistakes the individual-level intuition for the population-level result. You
  should address this directly by framing the contribution carefully. I'd avoid calling it a "new
  mechanism of superspreading" in the abstract or the first time you mention it --- that invites the
  dismissive reaction. Instead, frame it as: "We show that the interaction between infectiousness
  timing and time-varying contacts generates overdispersion in secondary infection counts that is not
   captured by existing frameworks for individual variation in transmission." Then, when you unpack
  it, emphasize the three non-obvious aspects above. The word "mechanism" is fine in the body of the
  paper, but lead with the consequence (overdispersion in a population with no individual-level
  heterogeneity), not the label (new mechanism).

  Concretely, in the Nature draft, I'd rewrite the opening of the superspreading paragraph from:

  "Here, we identify a third mechanism: when infectiousness is concentrated in a narrow temporal
  window, the overlap between that window and the contact process becomes a matter of chance."

  to something like:

  "We show that overdispersion in secondary infection counts can arise even in populations with no
  variation in biological infectiousness or average contact rates --- a result that cannot be
  explained by existing frameworks for superspreading. The mechanism is the stochastic interaction
  between a person's infectiousness window and their contact process: when infectiousness is
  concentrated in a narrow window, the overlap with high- or low-contact periods becomes a matter of
  chance, generating variation in nu_i that is not attributable to any persistent individual
  characteristic and therefore cannot be targeted by identifying high-risk individuals."

  This version leads with the surprising result (overdispersion without heterogeneity) rather than
  the mechanism, and immediately explains why it matters (you can't target it). The "duh" objection
  has less purchase when the framing is "here's a result your models can't produce" rather than
  "here's a mechanism."

  Does this tank Nature/Science?

  No, I don't think either concern is disqualifying. The contact variation issue is addressable with
  a paragraph and some citations. The framing issue is addressable with careful writing. Neither is a
   fundamental flaw in the science.

  What would concern me more at Nature is the empirical estimation section --- the fact that your
  strongest results (measles and plague near psi = 0) are the ones most susceptible to the biases you
   acknowledge. If a referee pushes hard on "your low-psi estimates could just be ascertainment bias
  toward event-based transmission clusters," you need a response. One option is to frame the
  empirical estimates as hypothesis-generating rather than definitive: "these estimates motivate
  further investigation with purpose-collected data" rather than "we have shown that measles has psi
  near 0." Another is to add a sensitivity analysis showing that your estimates are robust to
  dropping the largest clusters, which would partially address the ascertainment concern. That's a
  relatively small addition to the analysis that could meaningfully strengthen the paper.