\section{Supplementary Materials}
\subsection{Supplementary Methods}

\subsubsection{Deriving population-level quantities from individual-level components}

\subsubsection{Simulation approach}

We used stochastic, individual-based epidemic simulations to investigate the impact of individual-level infectiousness timing on epidemic dynamics. The simulation model has a modular design: an infection attempt generator produces a set of transmission attempts for each infected individual, and a population-level simulation script processes these attempts to propagate the epidemic. We implemented two simulation variants: a finite-population model with susceptible depletion, and an infinite-population model where every infection attempt succeeds. Both are implemented in {\bf \texttt{R}} (version 4.5.0), with performance-critical inner loops reimplemented in C++ via {\bf \texttt{Rcpp}} for speed.

\paragraph{Infection attempt generation.}
For each newly infected individual $i$ (with infection time $t_i$), the infection attempt generator proceeds as follows:
\begin{enumerate}
	\item \textbf{Draw the number of infection attempts.} The number of secondary infection attempts $\chi_i$ is drawn from a $\text{Poisson}(b_i)$ distribution, where $b_i$ is the individual's biological infectiousness. Throughout, we set $b_i \equiv R_0$ for all individuals, so that variation in realized transmission arises solely from infectiousness timing and contact dynamics.
	\item \textbf{Draw the latent period.} A latent period $l_i$ is drawn from a $\text{Gamma}((1-\psi)\alpha, \beta)$ distribution, representing the delay between infection and the onset of transmissibility. When $\psi = 1$, the latent period is zero; when $\psi = 0$, it accounts for the full variance of the generation interval.
	\item \textbf{Draw the timing of each infection attempt.} For each attempt $j = 1, \ldots, \chi_i$, an independent jitter $\epsilon_j$ is drawn from a $\text{Gamma}(\psi\alpha, \beta)$ distribution. The time of infection attempt $j$ relative to the index case's infection time is $\tau_{ij} = l_i + \epsilon_j$. When $\psi = 0$, $\epsilon_j = 0$ and all attempts occur simultaneously at time $l_i$; when $\psi = 1$, the $\epsilon_j$ are independent draws from the full $\text{Gamma}(\alpha, \beta)$ generation interval distribution.
	\item \textbf{Thin for time-varying contacts (if applicable).} When a time-varying contact process $c_i(t)$ is active, each infection attempt at calendar time $t_i + \tau_{ij}$ is accepted with probability $c_i(t_i + \tau_{ij}) / c_{\max}$, where $c_{\max}$ is the supremum of the contact process, using Poisson thinning. For the periodic contact model (Eq.~\ref{eq:periodic_contacts}), all individuals share the same deterministic contact function $c_i(t) = 1 - c_{\text{amp}} \cos(2\pi t / c_{\text{per}})$, and $c_{\max} = 1 + c_{\text{amp}}$. For the stochastic contact model, each individual's piecewise-constant contact trajectory is generated by drawing switching times from a Poisson process with rate $\lambda$ and contact levels from a $\text{Gamma}(\sigma, \sigma)$ distribution; the contact trajectory is generated over an interval $[0, 5\bar{g}]$ to ensure enough contact levels are drawn to accommodate all infection attempts. Thinning is performed against the realized maximum contact level over that individual's trajectory.
	\item \textbf{Thin for detect-and-isolate interventions (if applicable).} When a detect-and-isolate intervention is active, an individual-specific detection time $D_i$ is drawn (relative to peak infectiousness $l_i$) according to the detection mechanism in effect (symptom-based or screening-based). The individual participates with probability $\rho$. If participating, all infection attempts occurring after detection ($\tau_{ij} > t_i + D_i$) are removed with probability $\eta$, representing the effectiveness of post-detection isolation.
\end{enumerate}

\paragraph{Finite-population simulation.}
The finite-population simulation tracks susceptible depletion in a population of size $N$. It proceeds as follows:
\begin{enumerate}
	\item \textbf{Initialization.} A single index case is infected at time $t = 0$. All other $N-1$ individuals are susceptible.
	\item \textbf{Event scheduling.} Each infection attempt is stored as a pending event with its scheduled calendar time $t_i + \tau_{ij}$. Events are maintained in a priority queue, ensuring that the earliest pending event is always processed next.
	\item \textbf{Event processing.} At each step, the earliest event is removed from the queue. A target individual is selected uniformly at random from the population ($N - 1$ individuals excluding the infector). If the target is susceptible, infection occurs: the target is marked as infected, and its own infection attempts are generated and queued. If the target is already infected or recovered, the attempt fails (``wasted'' on a non-susceptible).
	\item \textbf{Termination.} The simulation terminates when the event queue is empty (epidemic extinction) or when a stopping criterion is met (\eg, a specified number of infections or calendar time is reached).
\end{enumerate}

\paragraph{Infinite-population simulation.}
The infinite-population simulation is a special case in which every infection attempt succeeds (no susceptible depletion). This is equivalent to a branching process approximation of the early epidemic and is useful for studying early-epidemic dynamics without the confounding effect of herd immunity. The implementation is identical to the finite-population version except that step 3 is simplified: every queued event produces a new infection, with no target selection or susceptibility check.

\paragraph{Default parameters.}
Unless otherwise stated, simulations used a population size of $N = 10{,}000$, with $n_{\text{sim}} = 5{,}000$ replicate simulations per scenario. Epidemic establishment was defined as reaching $0.05 N = 500$ cumulative infections. To cover a range of infectiousness profile shapes, we considered $\psi \in \{0, 0.2, 0.5, 0.8, 1\}$.

\paragraph{Benchmark pathogens.}
For illustration, we focused on three sets of epidemiological parameters reflecting published estimates of $R_0$ and $g(\tau)$ for influenza, SARS-CoV-2 (omicron), and measles. The generation interval distribution $g(\tau) \sim \text{Gamma}(\alpha, \beta)$ was parameterized using the mean generation time $\bar{g}$ and variance $\text{Var}[g]$ reported in the literature, with $\alpha = \bar{g}^2 / \text{Var}[g]$ and $\beta = \bar{g} / \text{Var}[g]$.

\begin{table}[h]
\centering
\begin{tabular}{lccccc}
\hline
Pathogen & $R_0$ & $\bar{g}$ (days) & $\text{Var}[g]$ (days$^2$) & $\alpha$ & $\beta$ \\
\hline
Influenza & 2 & 3.2 & 4.41 & 2.32 & 0.73 \\
SARS-CoV-2 (omicron) & 6 & 7.05 & 20.8 & 2.39 & 0.34 \\
Measles & 12 & 12.2 & 13.10 & 11.36 & 0.93 \\
\hline
\end{tabular}
\caption{Benchmark pathogen parameters used in simulations.}
\label{tab:benchmark_pathogens}
\end{table}



\clearpage
\subsubsection{Overdispersion in daily infection counts from punctuated infectiousness profiles}
\label{sec:supp_daily_overdispersion}

Here we derive the index of dispersion of the daily infection count attributable to a single infector, show that it is monotonically decreasing in the punctuation parameter $\psi$, and show that the overdispersion aggregates to the population level.

\paragraph{Setup.}
Fix an index case $i$, infected at time $t_i$, who generates $\chi_i \sim \text{Poisson}(R_0)$ total secondary infections. Under the Gamma time-shift model, the generation interval for each secondary infection is $\tau = l_i + \epsilon$, where $l_i \sim \text{Gamma}((1-\psi)\alpha, \beta)$ is a shared latent period and $\epsilon \sim \text{Gamma}(\psi\alpha, \beta)$ is an independent individual jitter. Let $N_i(\tau)$ be the number of person $i$'s secondary infections that fall on day-since-infection $[\tau, \tau+1)$, and define the random landing probability
\begin{equation}
	Q_i(\tau) = \int_\tau^{\tau+1} f_\epsilon(u - l_i) \, du
\end{equation}
where $f_\epsilon$ is the $\text{Gamma}(\psi\alpha, \beta)$ density. The randomness in $Q_i(\tau)$ arises from the draw of $l_i$.

\paragraph{Mean of $Q_i(\tau)$.}
Taking the expectation over $l_i$ and applying Fubini's theorem:
\begin{equation}
	E[Q_i(\tau)] = \int_\tau^{\tau+1} E_{l_i}[f_\epsilon(u - l_i)] \, du = \int_\tau^{\tau+1} (f_l * f_\epsilon)(u) \, du = \int_\tau^{\tau+1} g(u) \, du \equiv q(\tau)
\end{equation}
since $f_l * f_\epsilon = g$ by construction. Thus $E[Q_i(\tau)]$ depends only on the population-level generation interval distribution and is independent of $\psi$.

\paragraph{Index of dispersion of $N_i(\tau)$.}
Conditional on $Q_i(\tau)$, the $\chi_i$ offspring are independently assigned to day $[\tau, \tau+1)$ with probability $Q_i(\tau)$. By the Poisson thinning property (marginalizing over $\chi_i$):
\begin{equation}
	N_i(\tau) \mid Q_i(\tau) \sim \text{Poisson}(R_0 \, Q_i(\tau))
\end{equation}
Applying the law of total variance:
\begin{align}
	\text{Var}[N_i(\tau)] &= E[\text{Var}[N_i(\tau) \mid Q_i(\tau)]] + \text{Var}[E[N_i(\tau) \mid Q_i(\tau)]] \nonumber \\
	&= R_0 \, E[Q_i(\tau)] + R_0^2 \, \text{Var}[Q_i(\tau)]
\end{align}
Since $E[N_i(\tau)] = R_0 \, E[Q_i(\tau)] = R_0 \, q(\tau)$, the index of dispersion is
\begin{equation}
	\text{ID}[N_i(\tau)] = \frac{\text{Var}[N_i(\tau)]}{E[N_i(\tau)]} = 1 + R_0 \, \frac{\text{Var}[Q_i(\tau)]}{E[Q_i(\tau)]}
	\label{eq:ID_individual}
\end{equation}

\paragraph{Extreme cases.}
In the smooth case ($\psi = 1$), $f_\epsilon = g$ and $l_i = 0$ deterministically, so $Q_i(\tau) = q(\tau)$ is constant across individuals. Thus $\text{Var}[Q_i(\tau)] = 0$ and $\text{ID}[N_i(\tau)] = 1$.

In the spike case ($\psi = 0$), $f_\epsilon$ is a point mass at zero. Then $Q_i(\tau) = \mathbf{1}(l_i \in [\tau, \tau+1))$, which is Bernoulli with parameter $q(\tau)$. Thus $\text{Var}[Q_i(\tau)] = q(\tau)(1 - q(\tau))$ and
\begin{equation}
	\text{ID}[N_i(\tau)] = 1 + R_0(1 - q(\tau))
\end{equation}

% \paragraph{Monotonicity of $\text{Var}[Q_i(\tau)]$ in $\psi$.}
% We show that for $\psi_1 < \psi_2$, $\text{Var}[Q_i^{(\psi_1)}(\tau)] \geq \text{Var}[Q_i^{(\psi_2)}(\tau)]$, where the superscript makes the $\psi$-dependence explicit. The proof uses a coupling argument.

% Since Gamma distributions with the same rate parameter are additive, and $(1-\psi_1)\alpha > (1-\psi_2)\alpha$, we can write
% \begin{equation}
% 	L_{\psi_1} \stackrel{d}{=} L_{\psi_2} + \Delta
% \end{equation}
% where $L_{\psi_j} \sim \text{Gamma}((1-\psi_j)\alpha, \beta)$ and $\Delta \sim \text{Gamma}((\psi_2 - \psi_1)\alpha, \beta)$ is independent of $L_{\psi_2}$.

% Under this coupling, we establish a smoothing relationship. Let $Y = \mathbf{1}(\tau_{\text{gen}} \in [\tau, \tau+1))$ where $\tau_{\text{gen}} = L + \epsilon$ is a secondary infection's generation interval. Then $Q_i^{(\psi)}(\tau) = P(\tau_{\text{gen}} \in [\tau, \tau+1) \mid L_\psi) = E[Y \mid L_\psi]$. By the tower property:
% \begin{align}
% 	Q_i^{(\psi_2)}(\tau) &= E[Y \mid L_{\psi_2}] = E\!\left[E[Y \mid L_{\psi_2}, \Delta] \mid L_{\psi_2}\right] \nonumber \\
% 	&= E\!\left[E[Y \mid L_{\psi_1}] \mid L_{\psi_2}\right] = E\!\left[Q_i^{(\psi_1)}(\tau) \mid L_{\psi_2}\right]
% \end{align}
% where the third equality uses the fact that, given $(L_{\psi_2}, \Delta)$, we know $L_{\psi_1} = L_{\psi_2} + \Delta$, and $Y$ depends on $L_{\psi_1}$ and $\epsilon_{\psi_1}$ only, with $\epsilon_{\psi_1}$ independent of $(L_{\psi_2}, \Delta)$.

% Thus $Q_i^{(\psi_2)}(\tau)$ is a conditional expectation of $Q_i^{(\psi_1)}(\tau)$. Applying the law of total variance to $Q_i^{(\psi_1)}(\tau)$, conditioning on $L_{\psi_2}$:
% \begin{equation}
% 	\text{Var}[Q_i^{(\psi_1)}(\tau)] = \text{Var}[Q_i^{(\psi_2)}(\tau)] + E[\text{Var}[Q_i^{(\psi_1)}(\tau) \mid L_{\psi_2}]] \geq \text{Var}[Q_i^{(\psi_2)}(\tau)]
% \end{equation}
% which establishes the claimed monotonicity.

% \paragraph{Aggregation to population-level daily counts.}
% On calendar day $t$, the total number of new infections is $I(t) = \sum_i N_i(t - t_i)$, summing over all previously infected individuals $i$. Conditional on the epidemic history $\{t_i\}$, the $N_i(t - t_i)$ are independent (each infector generates offspring independently). Writing $q_i = q(t - t_i)$ for the population-level landing probability specific to infector $i$:
% \begin{equation}
% 	\text{Var}[I(t) \mid \{t_i\}] = \sum_i \text{Var}[N_i(t - t_i)]
% \end{equation}
% In the smooth case ($\psi = 1$), each $N_i(t-t_i) \sim \text{Poisson}(R_0 q_i)$, so $I(t) \mid \{t_i\} \sim \text{Poisson}(\sum_i R_0 q_i)$ and $\text{ID}[I(t) \mid \{t_i\}] = 1$.

% In the spike case ($\psi = 0$), using $\text{Var}[N_i(t-t_i)] = R_0 q_i + R_0^2 q_i(1-q_i)$:
% \begin{align}
% 	\text{ID}[I(t) \mid \{t_i\}] &= \frac{\sum_i \left[R_0 q_i + R_0^2 q_i(1-q_i)\right]}{\sum_i R_0 q_i} = 1 + R_0\!\left(1 - \frac{\sum_i q_i^2}{\sum_i q_i}\right) = 1 + R_0(1 - \bar{q}_w)
% 	\label{eq:ID_aggregate}
% \end{align}
% where $\bar{q}_w = \sum_i q_i^2 / \sum_i q_i$ is the intensity-weighted mean landing probability. Since $q_i \in [0,1]$, we have $q_i^2 \leq q_i$ for all $i$, so $\bar{q}_w \leq 1$. In a growing epidemic, infectors have a spread of ages since infection, so the $q_i$ values are generally well below 1, making $\bar{q}_w$ small and the overdispersion substantial. The aggregate index of dispersion is bounded: $\text{ID}[I(t) \mid \{t_i\}] \in [1, 1 + R_0]$, with ID $\to 1$ when the generation interval concentrates on a single day and ID $\to 1 + R_0$ when it is spread over many days.

\clearpage


\subsection{Supplementary Figures}